CF Registry Analysis

### 2023-03-09 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])),
            Age = Age_YrEnd[1],.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```

@@

>>>>

# Incidence and Prevalence

```{r}
# All data
encounter <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])),
            Age = Age_YrEnd[1],.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```

@@

# ROC

```{r}
# Prepare data for ROC
df$cfrd_ogtt <- factor(df$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
cfrd <- df %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ever = max(cfrd_ogtt),
    time_from_diag = ReviewYear - first_cfrd
  ) %>%
  filter(cfrd_ever == "Yes", time_from_diag <= 0, time_from_diag >= -1) %>%
  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
  drop_na()
non_cfrd <- df %>%
  group_by(eDWID) %>%
  mutate(cfrd_ever = max(cfrd_ogtt)) %>%
  filter(cfrd_ever == "No") %>%
  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
  drop_na()
```

```{r}
#| label: fig-roc
#| fig-cap: "ROC Curve"
r <- roc(
  cases = cfrd$hba1c, controls = non_cfrd$hba1c, quiet = T, plot = T, auc = T, ci = T,
  print.auc = TRUE, show.thres = TRUE
)
```

Because we do not have exact information regarding HbA1c timing (we only have encounter number within year), the annualized HbA1c from the year of diagnosis or from one year prior to diagnosis was used in the ROC analysis. For those with HbA1c measures the year of diagnosis and one year prior, the value from year of diagnosis was preferred. For those never diagnosed with CFRD, the most recent HbA1c was used. There were a total of `r nrow(cfrd)` HbA1c measures for people with CFRD and `r nrow(non_cfrd)` measures for those without.

The AUC for this ROC curve was 0.61 (see @fig-roc).

```{r}
#| label: tbl-roc_thresh
#| tbl-cap: "Best ROC Threshold (Youden Index)"
kable(coords(r, "best"), digits = 3)
```
